---
title: "Functional Analysis of Gene Lists - Exercises"
author: "Mireia Ferrer, Ricardo Gonzalo i Alex Sanchez"
date: "16/12/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This document lists some exercises that can be done with R/Bioconductor or using standalone or web tools.

# Preparing a gene list from a "top table"

- Obtain a file from where you can extract some genes following a "reasonable" criteria.
- Select 
    - The list of genes with a p-value less than 0.01/0.05
    - The list of genes with an adjusted p-value less than 0.05/0.15
    - The list of genes with an adjusted p-value less than 0.05/0.015 and an absolute (logFC) greater than 1.5
    
```{r}
ExpressAndTop_AvsB <- read.csv2("ExpressAndTop_AvsB.csv2")
library(dplyr)
topTabPvalLT01 <- filter(ExpressAndTop_AvsB, P.Value < 0.01)
entrezsPvalLT01 <- select(topTabPvalLT01, EntrezsA)
entrezsPvalLT01Bis <- ExpressAndTop_AvsB %>% 
  filter( P.Value < 0.01) %>% select(EntrezsA)
entrezsAdjPvalLT05 <- ExpressAndTop_AvsB %>% 
  filter( adj.P.Val < 0.05) %>% select(EntrezsA)
entrezsAdjPvalLT05logFCGT1.5 <- ExpressAndTop_AvsB %>% 
  filter( adj.P.Val < 0.05, abs(logFC) > 1.5) %>% select(EntrezsA)

```

- Repeat the exercise with a file of your own (for example the output of the analysis of the `airwayy` use case).

# Navigating the Gene Ontology

- The GO can be used to search for annotations, gene (products) or GO Terms

- Go to the GO web site and search for

    - Annotations related to psoriasis
    - Genes related with term psoriasis
    
# Enrichment analysis of a gene list

- Select a list of genes (ex. genes with an adjusted p-value < 0.05")
- Paste them into the GO Enrichment Analysis box in the GO web site

```{r}
write.table(entrezsAdjPvalLT05logFCGT1.5, file="genes.txt", sep="\t", 
            row.names=FALSE, col.names = FALSE)
```

- Repeat the exercise with a gene list of your own (for example a list that you have prepared with the output of the analysis of the `airwayy` use case).
